% !TEX root = ../multivar-notes.tex
\lessondate{January 8, 2019}

\Sec{Subsets $\R^n$ -- Point-set topology}

\begin{defn}
  Let $x\in \R^n$ and $\epsilon > 0$. Then $B_\epsilon (x) = \Set{y\in \R^n \mid |x-y|<\epsilon}$. This is called an $\epsilon$-neighborhood of $x$.
\end{defn}
\begin{defn}
  Let $X\subseteq \R^n$ and $x\in R^n$. Then x is called
  \begin{itemize}
    \item interior point of $X$ if $\exists \epsilon > 0$ such that $B_\epsilon (x) \subseteq X$.
    \item boundary point of $X$ if $\forall \epsilon > 0$, $B_\epsilon (x) \cap X \neq \emptyset$ and $B_\epsilon (x) \cap X^c \neq \emptyset$
    \item exterior point of $X$ if it is an interior point of $X^c$
  \end{itemize}
  Notation: $\mathring{X} = \text{interior of }X = \text{set of all interior point of }X$.
  $\delta X = \text{boundary of }X = \text{set of all boundary points of }X$
\end{defn}
\begin{defn}
  $X$ is called \ul{open} if it only consists of interior points. ($X=\mathring{X}$) \\
  $X$ is called \ul{closed} if its complement is open. \\
  $\implies X$ is open if it contains none of its boundary points. \\
  $\implies X$ is closed if it contains all of its boundary points \\
\end{defn}

\exercise{1.5.1, book p.101}
For each of the following subsets, state whether it is open or closed (or both or neither), and say why.

\begin{enumerate}[a.]
	\item $\Set{x\in \R \mid 0<x \leq 1}$ as a subset of $\R$ \\
	\answer{Neither.} $1$ is not an interior point of this set and $0$ is not an interior point of the complement of the set.
	\item $\Set{\Point{x \\ y}\in \R^2 \mid \sqrt{x^2+y^2}< 1}$ as a subset of $\R^2$ \\
	\answer{Open.} The unit circle (which is the boundary) is not contained within the set.
	\item the interval $(0,1]$ as a subset of $\R$ \\
	\answer{Neither.} Similar to a.
	\item $\Set{\Point{x \\ y}\in \R^2 \mid \sqrt{x^2+y^2}\leq 1}$ as a subset of $\R^2$ \\
	\answer{Closed.} This is the unit circle on the plane, and the boundary point set $x^2+y^2=1$ is wholly contained within this subset. 
	
	\item $\{x\in\R\mid 0\leq x\leq 1\}$ as a subset of $\R$. \\
	\answer{Closed.} Both boundary points, $0$ and $1$ are contained within this set. 
	
	\item $\{(x,y,z)\in\mathbb R^3 \mid \sqrt{x^2+y^2+z^2}=1 \text{ and }x,y,z\neq 0\}$ as a subset of $\R^3$. \\
	\answer{Closed.} This constitutes its own boundary points, where every $(x,y,z)$'s nbhd has intersections with both the set and the complement of the set. It is the unit spherical shell in 3-dimensions. 
	
	\item the empty set as a subset of $\R$. \\
	\answer{Both open and closed.} Its complement, the set of all real numbers, contains all of its boundary points (of which it has none) and contains none of its boundary points (of which it has none). 
\end{enumerate}


\exercise{1.5.2, book p.101}
For each of the following subsets, state whether it is open or closed (or both or neither), and say why.
\begin{enumerate}[a.]
\item $(x,y)$-plane in $\R^3$
\item $\R\subset \C$
\item the line $x=5$ in the $(x,y)$-plane
\item $(0,1)\subset \C$
\item $\R^n\subset \R^n$
\item the unit sphere in $\R^3$
\end{enumerate}

\exercise{1.5.5}
For each of the following subsets of $\R$ and $\R^2$, state whether it is open or closed (or both or neither), and prove it.
\begin{enumerate}[a.]
	\item $\Set{\Point{x \\ y} \in \R^2 \mid 1 < x^2+y^2 < 2}$ \\
	\item $\Set{\Point{x \\ y} \in \R^2 \mid xy\neq 0}$ \\
	\item $\Set{\Point{x \\ y} \in \R^2 \mid y=0}$
	\item $\Set{\mathbb{Q} \subset \R}\qquad$ (the rational numbers)
\end{enumerate}
